Outdated and License-Violating Code on Stack Overflow

### 2017-10-06 cloverflow-tse/cloverflow.tex
Added technical report updates.
Added latest results to the TSE draft.

\newcommand\FIXME[1]{\textbf{FIXME: #1}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
% We studied online code clones on the popular programming Q\&A
% website, Stack Overflow, and their issues on outdated code and
% software licensing violations. 
Online code clones are code fragments that are copied from software
projects or online sources to Stack Overflow as examples in
questions and answers.
% They can be discovered by using code clone
% detectors, yet they are inherently different from normal code clones
% in software projects in several aspects. 
Due to an absence of a checking mechanism after the code has been
copied to Stack Overflow, online code clones may suffer from
becoming outdated or violating the original software license.
% Moreover, we also found that majority of online code clones do not
% contain the license of their originals.
We present a study incorporating several state-of-the-art tools and
techniques to automatically extract, prioritise, and filter detected
online clone pairs between 144,064 Java code snippets on Stack
Overflow and 111 Java open source projects in the curated Qualitas
corpus. We discovered and analysed 32,533 non-trivial online clone
candidates. Our investigation of 3,636 manually confirmed clone
pairs revealed strong evidence that 112 clones have been copied from
a Qualitas project to Stack Overflow.  We found 64 of them (57\%) 
to be outdated and potentially harmful for reuse.  Furthermore, we
found 357 code snippets on Stack Overflow that violate the license
of their original software.
% These snippets are questionable for reuse.
% Our study is among the first to address the problem of
% outdated and license-violating code snippets on Stack Overflow.
\end{abstract}

@@

%The popularity of the Internet encourages tremendous amount of source code being shared online. 
Stack Overflow is a popular online programming community with 6.8
million users\footnote{Statistics as of February 25, 2017 from \url{http://stackexchange.com/sites}}. It allows programmers to ask questions and give answers
to programming problems. The website has found to be useful for
software
development~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014,
	Stolee2014,Subramanian2013,Diamantopoulos2015,Treude2016} and also
valuable for educational purposes~\cite{Nasehi2012}. On Stack
Overflow, each conversation contains a question and a list of
answers. The answers frequently contain at least one code snippet as
a solution to the question asked. We found that the code snippets are
usually not authored directly on the Stack Overflow website but copied
from another location. A snippet in an answer could be copied and modified
from a code snippet in the question, copied from the answerer's own
code or from other locations including open source software (OSS)
systems. The process of posting and answering questions on Stack
Overflow that involves reusing of source code, can be considered as
code cloning.

@@

In this study, we are interested in locating online code clones on
Stack Overflow that are cloned from open source projects and study
them on two aspects: \textbf{outdated code} and \textbf{license conflicts.}

\begin{figure*}
	\begin{lstlisting}
/* Code in Stack Overflow post ID 22315734 */         /* WritableComparator.java (2016-09-26) */
public int compare(byte[] b1,int s1,int l1, ...) {    public int compare(byte[] b1,int s1,int l1, ...) {
  try {                                                 try {
    buffer.reset(b1,s1,l1); /* parse key1 */              buffer.reset(b1,s1,l1); /* parse key1 */
    key1.readFields(buffer);                              key1.readFields(buffer);
    buffer.reset(b2,s2,l2); /* parse key2 */              buffer.reset(b2,s2,l2); /* parse key2 */
    key2.readFields(buffer);                              key2.readFields(buffer);
  } catch (IOException e) {                               $$@buffer.reset(null,0,0); /* clean up reference */@
    throw new RuntimeException(e);                      } catch (IOException e) {
  }                                                       throw new RuntimeException(e);
  return compare(key1,key2); /* compare them */         }
}                                                       return compare(key1, key2); /* compare them */
                                                      }
	\end{lstlisting}\vspace{-2ex}
	\caption{A motivating example of the two code fragments of
		{\small\texttt{WritableComparator.java}}. The one from the
		Stack Overflow post 22315734 (left) is outdated when compared to
		its latest version in the \textsf{hadoop} code base
		(right).}
	\label{fig:before-after}
\end{figure*}

A motivating example of outdated online code clones can be found in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{hadoop}\footnote{\url{http://stackoverflow.com/questions/22315734}}.
\Cref{fig:before-after} shows---on the left---a code snippet embedded as a part of the
accepted answer. The snippet shows how \textsf{hadoop}
implements the {\small{\texttt{compare}}} method in its
{\small{\texttt{WritableComparator}}} class. The code snippet on the
right shows another version of the same method, but at this time
extracted from the latest version (as of September 26, 2016) of
\textsf{hadoop}. We can see that they both are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added in
the latest version. The added line is intended for cleaning up the
reference in the {\small{\verb|buffer|}} variable. While this change
has already been introduced into the {\small{\texttt{compare}}} method
in the latest version of \textsf{hadoop}, the code example in Stack
Overflow post is still unchanged. This example shows that
inconsistencies between online code clones and their originals can
lead users of Stack Overflow to reuse outdated code.
%The code snippet on Stack Overflow can be outdated. %Since reusing source code from Stack Overflow is considered a common practice nowadays, the scale of online code cloning is more than intra or inter project clone in a local code bases.
%This is an emerging and challenging problem. Since studies in this area are still limited, we aim to gain more insight of the problem in this study. %We are interested to gain more insights of online code clone in this study.

While research has mostly focused on reusing code snippets \emph{from}
Stack Overflow (e.g.~\cite{Keivanloo2014,An2017,Yang2016}), fewer studies
have been conducted on finding the origins of code examples copied
\emph{to} Stack Overflow and problems of reusing them such as outdated
code and software licensing violations. This paper fills this gap and
makes the following primary contributions:
\begin{enumerate}
	\item \textbf{A manual study of online code clones:} 
	We used two clone detection tools to discover 315,786,118 similar code
	snippet pairs, followed by a manual classification of 3,636 candidate
	clone pairs between 144,064 Java code snippets obtained from Stack Overflow's
	accepted answers and 111 Java open source projects from the curated
	Qualitas corpus~\cite{QualitasCorpus}.
	\item \textbf{An investigation of outdated and license-violating
		online clones on Stack Overflow:} Our study shows that from the
	3,636 online clones, at least 671 likely have been copied from open
	source projects or external online sources to Stack Overflow,
	potentially violating software licenses. For 112 of them, we found
	evidence that they have been copied from a specific Qualitas project
	and that 64 of them are meanwhile outdated.
	% and questionable for being reused.
	\item \textbf{An online code clone oracle:} The 3,636 manually
	investigated and validated online clone pairs are available for
	download\footnote{https://to.be.disclosed.after.review} and can be
	used as a clone oracle.
\end{enumerate}

@@

We performed an empirical study of online code clones between Stack
Overflow and 111 Java open source projects to answer the following
research questions:
\begin{itemize}
	\item \textbf{RQ1 (Online code clones): }\textit{To what extent is source
		code cloned between Stack Overflow and open source projects?} We
	quantitatively measured the number of online code clones between Stack
	Overflow and open source projects to understand the scale of the
	problem. 
	\item \textbf{RQ2 (Patterns of online code clones): }\textit{Why do online
		code clones occur?} We categorised online clones into seven
	categories allowing insights to why online code clones are created.
	\item \textbf{RQ3 (Outdated online code clones): }\textit{Are
		online code clones up-to-date compared to their counterparts in the
		original projects?} We were interested in the outdated Stack
	Overflow code examples since users are potentially reusing
	them. 
	\item \textbf{RQ4 (Software licensing violation): }\textit{Do
		licensing conflicts occur between Stack Overflow clones and their
		originals?} We investigated whether the reuse of online code clones
	can cause software developers to violate licenses.
	\item \textbf{RQ5 (Developers' awareness): }\textit{Do Stack Overflow
		users aware of the outdated code and licensing problem when
		they answer a question on Stack Overflow?} 
	We surveyed 600 high reputation Stack Overflow users to study their
	awareness of the issues.
\end{itemize}

@@

%\subsection{Experimental Setup}
%\vspace{1ex}
To answer these four research questions, we perform an empirical study to
locate and study the online code clones between Stack Overflow and open source
projects. 
%
%\subsection{Methodology}
%
We designed our study in 6 phases as depicted in \Cref{fig:exp_framework} where
we build different data sets to answer each of our four research
questions.

@@

\textbf{Stack Overflow:} 
We extracted Java code snippets from a snapshot of a Stack Overflow
dump\footnote{\url{https://archive.org/details/stackexchange}} in January 2016.
% The archived dump has a size of 9 gigabytes.
The data dump is in XML format containing information of posts (questions and
answers) and we were interested in code snippets embedded in posts which were
located between {\small\texttt{<code>}...\texttt{</code>}} tags. A Stack
Overflow thread contains a question and several answers. An answer can also be
marked as \textbf{accepted answer} by the questioner if the solution fixes
his/her problem. We collected Java code snippets using two criteria. First, we
only focused on code snippets in accepted answers. We chose the snippets in
accepted answers because they actually solved the problems in the questions.
Moreover, they are always displayed just below the questions which makes them
more likely to be reused than other answers. Second, we were only interested in
code snippets of at least ten lines. Although the minimum clone size of six
lines are usual in clone detection~\cite{Bellon2007,Wang2013,Koschke2006},
snippets of at least ten lines are usually more interesting since most of
boiler-plate code of \texttt{equal} or \texttt{hashCode} methods are removed.
Each snippet was extracted from the dump and saved to a file. Moreover, we
filtered out irrelevant code snippets that were part of the accepted answers
but were not written in Java by using regular expressions and manual checking.
~\FIXME{should I mention I use PMD CPD to help here? More confusing?}.
Finally we obtained
72,365 Java code snippets containing 1,840,581 lines\footnote{Measured by cloc:
	\url{https://github.com/AlDanial/cloc}} of Java source code. The median of the
snippet size is 10.

@@

\begin{table}
	\centering
	\caption{Stack Overflow and Qualitas datasets}
	\label{tab:datasets}
	%\small
	\begin{tabular}{lrrr}
		\toprule
		Dataset & No. of files & SLOC & Median \\
		\midrule
		Stack Overflow & 72,365 & 1,840,581 & 10 \\ 
		Qualitas &  166,709 & 19,614,083 & 60,667 \\ 
		\bottomrule
	\end{tabular} 
\end{table}

\textbf{Clone Detection Tools: } We use clone detection to discover online code
clones. %\textbf{Clone Detection and Configuration: }
There are a number of restrictions in terms of choosing the clone detection
tools for this study. The main restriction is due to nature of code snippets
posted on Stack Overflow, as most of them are incomplete Java classes or
methods. Hence, a detector must be flexible enough to process code snippets that
are not compilable or not complete blocks. Moreover, since the amount of code
that has to be processed is in a scale of millions line of code (as shown in
\Cref{tab:datasets}), a clone detector must be scalable enough to report clones
in a reasonable amount of time. We have tried 7 state-of-the-art clone detectors
including Simian~\cite{simian}, SourcererCC~\cite{Sajnani2016},
NiCad~\cite{Cordy,Roy2008}, CCFinder~\cite{Kamiya2002}, iClones~\cite{Gode2009},
DECKARD~\cite{Jiang2007a}, and PMD-CPD~\cite{pmd-cpd} against the Stack Overflow
and Qualitas datasets. NiCad failed to parse 44,960 Stack Overflow snippets
while PMD CPD failed to complete the execution due to lexical errors. iClones
could complete its execution but skipped 367 snippets due to malformed blocks in
Stack Overflow data set. Although Simian, SourcererCC, CCFinder, and DECKARD
could successfully report clones, we decided to choose only Simian and
SourcererCC due to their fast detection speed.

@@

\subsection{Phase 2: Clone Merging}
As there are in total 315,786,118 clone pairs reported, it is
infeasible for humans to manually validate them all. Instead of
investigating a random sample, we adopted the idea of \textbf{clone
	agreement} which has been used in clone research
studies~\cite{Funaro2010, Wang2013,cr2016ssbse} in situations where a
clone oracle is missing or impossible to establish. Clone pairs agreed
by multiple clone detection tools have a higher likelihood to be real
clones~\cite{cr2016ssbse}. By using this agreement-based approach, we
reduced the number of clone candidates for manual investigation to the
ones agreed by multiple tools. To find agreement between two clone
pairs reported by two different tools, we used the clone pair matching
metric proposed by Bellon et al.~\cite{Bellon2007}. Two clone pairs
which have a large enough number of overlapping lines can be
categorised as either a good-match or an ok-match pair (denoted as
\textit{good} and \textit{ok} respectively) with a confidence value
between 0 and 1. A \textit{good} clone pair has stronger agreement
than an \textit{ok} pair. We follow Bellon's original definitions of
good- and ok-match~\cite{Bellon2007}, which are based on how much two
clone fragments \textit{CF} overlap\footnote{The overlap value can be
	considered as Jaccard similarity.}  and how much of them is
contained in each other:
%\squeezeup
\begin{displaymath}
overlap(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1) \cup lines(\textrm{\textit{CF}}_2)|} 
\end{displaymath}
% \squeezeup
\begin{displaymath}
contained(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1)|}
\end{displaymath}
\noindent%
A clone pair \textit{CP} is formed by two clone fragments
\textit{CF$_1$} and \textit{CF$_2$}, i.e.~\textit{CP} =
(\textit{CF$_1$}, \textit{CF$_2$}), and the \textit{good-value} and
the \textit{ok-value} of two clone pairs are defined as

%\squeezeup
\begin{align*}
good(\textrm{\textit{CP}}_1, \textrm{\textit{CP}}_2) = min(overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1), \\ overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2))
\end{align*}
\begin{align*}
ok(\textrm{\textit{CP}}_1,\textrm{\textit{CP}}_2) = min(max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1),~~ \\ contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1)),~
\\ max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2),~~ \\contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2)))
\end{align*}

>>>>

\newcommand\FIXME[1]{{\color{red}\textbf{FIXME: #1}}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
% We studied online code clones on the popular programming Q\&A
% website, Stack Overflow, and their issues on outdated code and
% software licensing violations. 
\FIXME{Outdated.~Revise the last.}~Online code clones are code fragments that are copied from software
projects or online sources to Stack Overflow as examples in
questions and answers.
% They can be discovered by using code clone
% detectors, yet they are inherently different from normal code clones
% in software projects in several aspects. 
Due to an absence of a checking mechanism after the code has been
copied to Stack Overflow, online code clones may suffer from
becoming outdated or violating the original software license.
% Moreover, we also found that majority of online code clones do not
% contain the license of their originals.
We present a study incorporating several state-of-the-art tools and
techniques to automatically extract, prioritise, and filter detected
online clone pairs between 144,064 Java code snippets on Stack
Overflow and 111 Java open source projects in the curated Qualitas
corpus. We discovered and analysed 32,533 non-trivial online clone
candidates. Our investigation of 3,636 manually confirmed clone
pairs revealed strong evidence that 112 clones have been copied from
a Qualitas project to Stack Overflow.  We found 64 of them (57\%) 
to be outdated and potentially harmful for reuse.  Furthermore, we
found 357 code snippets on Stack Overflow that violate the license
of their original software.
% These snippets are questionable for reuse.
% Our study is among the first to address the problem of
% outdated and license-violating code snippets on Stack Overflow.
\end{abstract}

@@

%The popularity of the Internet encourages tremendous amount of source code being shared online. 
Stack Overflow is a popular online programming community with 7.6
million users, 14 million questions, and 23 million answers\footnote{Data as of 21
	August 2017 from~\url{https://stackexchange.com/sites}}. It allows programmers to ask questions and give answers
to programming problems. The website has found to be useful for
software
development~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014,
	Stolee2014,Subramanian2013,Diamantopoulos2015,Treude2016} and also
valuable for educational purposes~\cite{Nasehi2012}. On Stack
Overflow, each conversation contains a question and a list of
answers. The answers frequently contain at least one code snippet as
a solution to the question asked. We found that the code snippets are
usually not authored directly on the Stack Overflow website but copied
from another location. A snippet in an answer could be copied and modified
from a code snippet in the question, copied from the answerer's own
code or from other locations including open source software (OSS)
systems. The process of posting and answering questions on Stack
Overflow that involves reusing of source code, can be considered as
code cloning.

@@

The findings from our online survey of Stack Overflow answerers show that code
reusing activity occurs on Stack Overflow, just like in software. Stack Overflow
answerers occasionally clone code from other locations, such as their personal
projects, company projects, and open source projects, to Stack Overflow answers
as a solution or to complement their answer~\FIXME{put a precise number here}.
One participants says

\begin{quote}\textit{The real issue is less about the amount the code snippets
		on SO than it is about the staggeringly high number of software
		``professionals'' that mindlessly use them without understanding what they're
		copying, and the only slightly less high number of would-be professionals that
		post snippets with built-in security issues.  A related topic is beginners who
		post (at times dangerously) misleading tutorials online on topics they actually
		know very little about. Think PHP/MySQL tutorials written 10+ years after
		\texttt{mysql\_*} functions were obsolete, or the recent regex tutorial that
		got posted the other day on HackerNew
		(\url{https://news.ycombinator.com/item?id=14846506}). They're also full of
		toxic code snippets.}\end{quote}

With the same analogy, in this study we are interested in mining Stack Overflow
posts, detecting online code clones, and revealing \textbf{toxic code snippets},
which means code snippets that are harmful for reuse.
Stack Overflow snippets can become toxic because they are outdated and violating
their original software license. Outdated code snippets can be harmful since
they may contain defects. Moreover, licensed code snippets that are cloned from
open source software projects to Stack Overflow without their respective license
can lead to legal issues if being reused.

\begin{figure*}
	\begin{lstlisting}
/* Code in Stack Overflow post ID 22315734 */         /* WritableComparator.java (2017-10-03) */
public int compare(byte[] b1,int s1,int l1, ...) {    public int compare(byte[] b1,int s1,int l1, ...) {
  try {                                                 try {
    buffer.reset(b1,s1,l1); /* parse key1 */              buffer.reset(b1,s1,l1); /* parse key1 */
    key1.readFields(buffer);                              key1.readFields(buffer);
    buffer.reset(b2,s2,l2); /* parse key2 */              buffer.reset(b2,s2,l2); /* parse key2 */
    key2.readFields(buffer);                              key2.readFields(buffer);
  } catch (IOException e) {                               $$@buffer.reset(null,0,0); /* clean up reference */@
    throw new RuntimeException(e);                      } catch (IOException e) {
  }                                                       throw new RuntimeException(e);
  return compare(key1,key2); /* compare them */         }
}                                                       return compare(key1, key2); /* compare them */
                                                      }
	\end{lstlisting}\vspace{-2ex}
	\caption{A motivating example of the two code fragments of
		{\small\texttt{WritableComparator.java}}. The one from the
		Stack Overflow post 22315734 (left) is outdated when compared to
		its latest version in the \textsf{hadoop} code base
		(right).}
	\label{fig:before-after}
\end{figure*}

We would like to motivate the readers to the problem of toxic code snippets
on Stack Overflow by giving two examples.

The first motivating example of outdated online code clones can be found in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{hadoop}\footnote{\url{http://stackoverflow.com/questions/22315734}}.
\Cref{fig:before-after} shows---on the left---a code snippet embedded as a part
of the accepted answer. The snippet shows how \textsf{hadoop} implements the
{\small{\texttt{compare}}} method in its {\small{\texttt{WritableComparator}}}
class. The code snippet on the right shows another version of the same method,
but at this time extracted from the latest version (as of October 3, 2017) of
\textsf{hadoop}. We can see that they both are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added on November
21, 2014. The added line is intended for cleaning up the reference in the
{\small{\verb|buffer|}} variable and avoid excess heap usage
(HADOOP-11323\footnote{https://issues.apache.org/jira/browse/HADOOP-11323}).
While this change has already been introduced into the
{\small{\texttt{compare}}} method for several years, the code example in Stack
Overflow post is still unchanged. 
%This example shows that inconsistencies
%between online code clones and their originals can lead users of Stack Overflow
%to reuse outdated code. 
In addition, although the original code snippet of
WritableComparator.java in Hadoop is subject to Apache license version 2.0, the
cloned code on Stack Overflow contains only a {\small{\texttt{compare}}}
method and misses its license statement on top of the file. There are two
potential issues for this. First, the code snippet may look like it is under
Stack Overflow's CC BY-SA 3.0 instead of its original Apache license. Second, if
the code snippet is copied and incorporated into another software project with
conflicting license, a legal issue may arise.

\begin{figure*}
	\begin{lstlisting}
/* Code in Stack Overflow post ID 801987 */             /* StringUtils.java (2017-10-03) */
public static String humanReadableInt(long number) {    public static String humanReadableInt(long number) {
  long absNumber = Math.abs(number);                      return TraditionalBinaryPrefix.long2String(number,"",1);
  double result = number;                               }
  String suffix = "";
  if (absNumber < 1024) {
  } else if (absNumber < 1024 * 1024) {
    result = number / 1024.0;
    suffix = "k";
  } else if (absNumber < 1024 * 1024 * 1024) {
    result = number / (1024.0 * 1024);
    suffix = "m";
  } else {
    result = number / (1024.0 * 1024 * 1024);
    suffix = "g";
  }
  return oneDecimal.format(result) + suffix;
}
	\end{lstlisting}\vspace{-2ex}
	\caption{A motivating example of the two code fragments of
		{\small\texttt{StringUtils.java}}. The one from the
		Stack Overflow post 801987 (left) is outdated when compared to
		its latest version in the \textsf{hadoop} code base
		(right). The outdated code has race conditions.}
	\label{fig:before-after_2}
\end{figure*}

The second motivating example of outdated online code clones with more
disrupting changes than the first one  can be found in an answer to a Stack
Overflow question regarding how to format files sizes in a human readable form.
\Cref{fig:before-after_2} shows---on the left---a code snippet to perform the
task from {\small{\texttt{StringUtils}}} class in \textsf{hadoop}. The code snippet on the
right shows another version of the same method, but at this time extracted from
the latest version of \textsf{hadoop}. We can see that
they both are totally different. The {\small{\texttt{humanReadableInt}}} method is
rewritten on February 5, 2013 to solve an issue of a race condition
(HADOOP-9252\footnote{https://issues.apache.org/jira/browse/HADOOP-9252}).

This two toxic code snippet has been posted on March 11, 2014 and April 9, 2009
respectively. They have already been viewed 259 and 2,886 times as the time of
writing this paper (October 3, 2017). Our calculation finds that there will be a
new viewer of the first toxic snippet in approximately every 5 days while the
second toxic snippet will be viewed almost every single day. Moreover, a
quick web search shows that the same {\small{\texttt{humanReadableInt}}}
method already reside in approximately 500 projects on GitHub. Considering the popularity of Stack
Overflow, which has more than 50 million developers visiting each
month\footnote{Data as of 21 August 2017 from: \url{https://stackoverflow.com}},
one toxic code snippets on Stack Overflow can spread and grow to hundred or
thousand copies within only a year or two.

%The code snippet on Stack Overflow can be outdated. %Since reusing source code from Stack Overflow is considered a common practice nowadays, the scale of online code cloning is more than intra or inter project clone in a local code bases.
%This is an emerging and challenging problem. Since studies in this area are still limited, we aim to gain more insight of the problem in this study. %We are interested to gain more insights of online code clone in this study.

While research has mostly focused on reusing code snippets \emph{from} Stack
Overflow (e.g.~\cite{Keivanloo2014,An2017,Yang2016}), fewer studies have been
conducted on finding the origins of code examples copied \emph{to} Stack
Overflow and the awareness of Stack Overflow developers in doing so. Finding the
origins of code examples gives insights to the problems of toxic code snippets
including outdated code and software licensing violations. It is probably more
important than studying the effects of reusing Stack Overflow code snippets
because it tackles the root cause of the problem. Moreover, this paper lays a
foundation to possibility of having an automatic technique to detect and report
such problems to developers in the future.

This paper fills this gap and makes the following primary contributions:

\begin{enumerate} 
	\item \textbf{A manual study of online code clones:} We used
	two clone detection tools to discover 315,786,118 similar code snippet pairs,
	followed by a manual classification of 3,636 candidate clone pairs between
	144,064 Java code snippets obtained from Stack Overflow's accepted answers and
	111 Java open source projects from the curated Qualitas
	corpus~\cite{QualitasCorpus}.

	\item \textbf{An investigation of outdated and
		license-violating online clones on Stack Overflow:} Our study shows that from
	the 3,636 online clones, at least 671 likely have been copied from open source
	projects or external online sources to Stack Overflow, potentially violating
	software licenses. For 112 of them, we found evidence that they have been copied
	from a specific Qualitas project and that 64 of them are meanwhile outdated. %
	% and questionable for being reused.

	\item \textbf{Awareness of Stack Overflow answerers to
	toxic code snippets:} We performed an online survey and collected answers from
	202 highly-ranked Stack Overflow developers. We found that there are~\FIXME{add
	the number here}\% of code snippets cloned from open source projects to Stack
	Overflow answers. While Stack Overflow answerers are aware of their outdated
	code snippets, 19\% of the participants report that they rarely or never fix
	the code. At least 98\% of the answerers never include software license in
	their snippets and 69\% never check for licensing conflicts with Stack
	Overflow's CC BY-SA 3.0 if they copy the code from other sources to Stack
	Overflow answers.

	\item \textbf{An online code
		clone oracle:} The 2,302 manually investigated and validated online clone pairs
	are available for download\footnote{https://to.be.disclosed.after.review} and
	can be used as a clone oracle. \end{enumerate}

@@

We performed an empirical study of online code clones between Stack
Overflow and 111 Java open source projects to answer the following
research questions:
\begin{itemize}
	\item \textbf{RQ1 (Online code clones): }\textit{To what extent is source
		code cloned between Stack Overflow and open source projects?} We
	quantitatively measured the number of online code clones between Stack
	Overflow and open source projects to understand the scale of the
	problem. 
	\item \textbf{RQ2 (Patterns of online code clones): }\textit{Why do online
		code clones occur?} We categorised online clones into seven
	categories allowing insights to why online code clones are created.
	\item \textbf{RQ3 (Outdated online code clones): }\textit{Are
		online code clones up-to-date compared to their counterparts in the
		original projects?} We were interested in the outdated Stack
	Overflow code examples since users are potentially reusing
	them. 
	\item \textbf{RQ4 (Software licensing violation): }\textit{Do
		licensing conflicts occur between Stack Overflow clones and their
		originals?} We investigated whether the reuse of online code clones
	can cause software developers to violate licenses.
	\item \textbf{RQ5 (Developers' awareness): }\textit{Do Stack Overflow
		users aware of the outdated code and licensing problem when
		they answer a question on Stack Overflow?} 
	We surveyed 607 high reputation Stack Overflow users to study their
	awareness of the issues.
\end{itemize}

@@

%\subsection{Experimental Setup}
%\vspace{1ex}
To answer these five research questions, we perform an empirical study to
locate and study the online code clones between Stack Overflow and open source
projects. 
%
%\subsection{Methodology}
%
We designed our study in 6 phases as depicted in \Cref{fig:exp_framework} where
we build different data sets to answer each of our four research
questions.

@@

\textbf{Stack Overflow:} 
We extracted Java code snippets from a snapshot of a Stack Overflow
dump\footnote{\url{https://archive.org/details/stackexchange}} in January 2016.
% The archived dump has a size of 9 gigabytes.
The data dump is in XML format containing information of posts (questions and
answers) and we were interested in code snippets embedded in posts which were
located between {\small\texttt{<code>}...\texttt{</code>}} tags. A Stack
Overflow thread contains a question and several answers. An answer can also be
marked as \textbf{accepted answer} by the questioner if the solution fixes
his/her problem. We collected Java code snippets using two criteria. First, we
only focused on code snippets in accepted answers. We chose the snippets in
accepted answers because they actually solved the problems in the questions.
Moreover, they are always displayed just below the questions which makes them
more likely to be reused than other answers. Second, we were only interested in
code snippets of at least ten lines. Although the minimum clone size of six
lines are usual in clone detection~\cite{Bellon2007,Wang2013,Koschke2006},
snippets of at least ten lines are usually more interesting since most of
boiler-plate code of \texttt{equal} or \texttt{hashCode} methods are removed.
Each snippet was extracted from the dump and saved to a file. Moreover, we
filtered out irrelevant code snippets that were part of the accepted answers
but were not written in Java by using regular expressions and manual checking.
~\FIXME{@Jens, should I mention I use PMD CPD to help here? More confusing?}.
Finally we obtained
72,365 Java code snippets containing 1,840,581 lines\footnote{Measured by cloc:
	\url{https://github.com/AlDanial/cloc}} of Java source code. The median of the
snippet size is 10.

@@

\begin{table}
	\centering
	\caption{Stack Overflow and Qualitas datasets}
	\label{tab:datasets}
	%\small
	\begin{tabular}{lrrr}
		\toprule
		Dataset & No. of files & SLOC & Median \\
		\midrule
		Stack Overflow & 72,365 & 1,840,581 & ?? \\ 
		Qualitas &  166,709 & 19,614,083 & 60,667 \\ 
		\bottomrule
	\end{tabular} 
\end{table}

\textbf{Clone Detection Tools: } We use clone detection to discover online code
clones. %\textbf{Clone Detection and Configuration: }
There are a number of restrictions in terms of choosing the clone detection
tools for this study. The main restriction is due to nature of code snippets
posted on Stack Overflow, as most of them are incomplete Java classes or
methods. Hence, a detector must be flexible enough to process code snippets that
are not compilable or not complete blocks. Moreover, since the amount of code
that has to be processed is in a scale of millions line of code (as shown in
\Cref{tab:datasets}), a clone detector must be scalable enough to report clones
in a reasonable amount of time. We have tried 7 state-of-the-art clone detectors
including Simian~\cite{simian}, SourcererCC~\cite{Sajnani2016},
NiCad~\cite{Cordy,Roy2008}, CCFinder~\cite{Kamiya2002}, iClones~\cite{Gode2009},
DECKARD~\cite{Jiang2007a}, and PMD-CPD~\cite{pmd-cpd} against the Stack Overflow
and Qualitas datasets. NiCad failed to parse 44,960 Stack Overflow snippets
while PMD CPD failed to complete the execution due to lexical errors. iClones
could complete its execution but skipped 367 snippets due to malformed blocks in
Stack Overflow data set. CCFinder reported 8 errors while processing the two data set. 
Although Simian, SourcererCC, and DECKARD
could successfully report clones, we decided to choose only Simian and
SourcererCC due to their fast detection speed.

@@

\subsection{Phase 2: Clone Merging}
As there are in total 315,786,118 clone pairs reported, it is infeasible for
humans to manually validate them all. Instead of investigating a random sample,
we adopted the idea of \textbf{clone agreement} which has been used in clone
research studies~\cite{Funaro2010, Wang2013,cr2016ssbse} in situations where a
clone oracle is missing or impossible to establish. Clone pairs agreed by
multiple clone detection tools have a higher likelihood to be real
clones~\cite{cr2016ssbse}. By using this agreement-based approach, we reduced
the number of clone candidates for manual investigation and avoid a threat to
internal validity of double counting the same clone pair twice by merging the
ones agreed by the two tools. To find agreement between two clone pairs
reported by two different tools, we used the clone pair matching metric proposed
by Bellon et al.~\cite{Bellon2007}. Two clone pairs which have a large enough
number of overlapping lines can be categorised as either a good-match or an
ok-match pair (denoted as \textit{good} and \textit{ok} respectively) with a
confidence value between 0 and 1. Although good-match has stronger
agreement than ok-match, we choose ok-match as our clone merging method because 
it does not take clone size into account which suit our results of clones from
two tools, Simian, SourcererCC, with two different detection methods, 
line-based and method-based.

\ We follow Bellon's original definitions of ok-match~\cite{Bellon2007}, which are based on how much two clone
fragments \textit{CF} is contained in each other:

%\squeezeup
%\begin{displaymath}
%overlap(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1) \cup lines(\textrm{\textit{CF}}_2)|} 
%\end{displaymath}
% \squeezeup
\begin{displaymath}
contained(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1)|}
\end{displaymath}
\noindent%
A clone pair \textit{CP} is formed by two clone fragments
\textit{CF$_1$} and \textit{CF$_2$}, i.e.~\textit{CP} =
(\textit{CF$_1$}, \textit{CF$_2$})
%, and the \textit{good-value} and
and the \textit{ok-value} of two clone pairs are defined as
